---
title: "week9"
author: "Lu Zhang"
date: "7/27/2019"
output:
  pdf_document: default
  html_document: default
---

# Sampling Distributions

## plotting system

### The base plotting system

 Scatterplot with loess curve
 
```{r}

data(airquality)
with(airquality, {
  plot(Temp, Ozone)
  lines(loess.smooth(Temp, Ozone))})

```

Add title

```{r}

data(cars)

## Create the plot / draw canvas
with(cars, plot(speed, dist))
 
## Add annotation
title("Speed vs. Stopping distance")


```


### The Lattice System

xyplot
bwplot


```{r}

library(lattice)

```

relationship between life expectancy and income and how that relationship varies by region

```{r}

state <- data.frame(state.x77, region = state.region)
xyplot(Life.Exp ~ Income | region, data = state, layout = c(4, 1))

```

### The ggplot2 System

```{r}

install.packages("ggplot2")

library(ggplot2)


```

qplot()
quickly get some data on the screen

```{r}

data(mpg)
qplot(cty, hwy, data = mpg)

```


## Graphics Devices

base plot

```{r}

## Make plot appear on screen device
with(faithful, plot(eruptions, waiting)) 

## Annotate with a title
title(main = "Old Faithful Geyser data") 


```

```{r}

## Open PDF device; create 'week9.pdf' in my working directory

pdf(file = "week9.pdf")  
# Create plot and send to a file (no plot appears on screen)
with(faithful, plot(eruptions, waiting))  

## Annotate plot; still nothing on screen
title(main = "Old Faithful Geyser data")  

## Close the PDF file device
dev.off()  

## the file 'week9.pdf' is in the directory

```

## Copying Plots

dev.copy()
  copy a plot from one device to another
  
```{r}

library(datasets)
 
## Create plot on screen device
with(faithful, plot(eruptions, waiting))  

## Add a main title
title(main = "Old Faithful Geyser data")  

## Copy my plot to a PNG file
dev.copy(png, file = "geyserplot.png")  

## Close the PNG device!
dev.off()  

```

