---
title: "week6"
author: "Lu Zhang"
date: "7/27/2019"
output:
  pdf_document: default
  html_document: default
---

# Data transformation

## Prerequisites: Packages

```{r setup, include=FALSE}

library(nycflights13)

```

```{r}
flights

```

## Filter rows with filter()

```{r}
sep6 <- filter(flights, month == 9, day == 6)

sep6
```


## Comparisons

###  >, >=, <, <=, != (not equal), and == (equal)

``` {r}

near(sqrt(3) ^ 2,  3)

near(sqrt(5) ^ 2,  3)

```

## Logical operators

### Boolean operators 

& is “and”, | is “or”, and ! is “not”


```{r}

filter(flights, month == 8 | month == 9)


```



```{r}

aug_sep <- filter(flights, month %in% c(8,9))

aug_sep

```



```{r}

filter(flights, !(arr_delay > 120 | dep_delay > 120))

```

## Missing values

determine if a value is missing

```{r}

x <- NA

is.na(x)

```


```{r}

df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)

# preserve missing values

filter(df, is.na(x) | x > 1)

```

## Arrange rows with arrange()

```{r}

arrange(flights, year, month, day)

```

re-order by a column in descending order using desc()

```{r}

arrange(flights, desc(dep_delay))


```

## Select columns with select()

Select columns by name

```{r}

select(flights, year, month, day, dep_time)

```

Select all columns between year and dep_time (inclusive)

```{r}

select(flights, year:dep_time)


```

Select all columns except those from distance to time_hour (inclusive)

```{r}

select(flights, -(year:day))

```

rename variables

```{r}

rename(flights, airtime = air_time )

```

## Add new variables with mutate()

```{R}

flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)

```

```{R}

mutate(flights_sml,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)


```

transmute(): only keep the new variables

```{R}

transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)

```

## Grouped summaries with summarise()

```{r}

by_day <- group_by(flights, year, month, day)

summarise(by_day, 
          delay = mean(dep_delay, 
                       na.rm = TRUE))

```

## Combining multiple operations with the pipe

explore the relationship between the distance and average delay for each carrier

```{r}

by_carrier <- group_by(flights, carrier)

delay <- summarise(by_carrier,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)

delay <- filter(delay, count > 50, carrier != "HNL")

ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)

```

## Missing values

removes the missing values prior to computation
na.rm = TRUE 


```{r}

flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))

```

missing values represent cancelled flights

```{r}

not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))

```

## Counts

a count (n())

or

a count of non-missing values
                            (sum(!is.na(x))


the planes (identified by their tail number) that have the highest average delays:

```{r}

delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )

ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)

```



```{r pressure, echo=FALSE}

delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )

ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)

plot(pressure)
```

only show delay more than 30 min

```{r pressure, echo=FALSE}

delays %>% 
  filter(n > 30) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
    geom_point(alpha = 1/10)

plot(pressure)
```

## Useful summary functions

```{r}

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0]) 
  )

```



```{r}

not_cancelled %>% 
  group_by(carrier) %>% 
  summarise(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))

```

Measures of rank: 

min(x), quantile(x, 0.25), max(x)

the total number of miles each carrier flew

```{r}

not_cancelled %>% 
  count(carrier, wt = distance)

```

