---
title: "week7"
author: "Lu Zhang"
date: "7/27/2019"
output: html_document
---

# Data Visualization

## Dataset

```{r}

install.packages("tidyverse")

library(tidyverse)

```

```{r}

mpg

```

```{r}

? mpg

```

## ggplot

plot mpg

ggplot()
geom_point() scatterplot

ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))

```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy))

```

##  Aesthetic mappings

color
alpha
shape
size

```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy, color = class))

```

```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy, shape = class))

```

## Facets

facet plot by a single variable
facet_wrap()

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy)) + 
  facet_wrap(~ class, nrow = 3)
```

facet_grid()

```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy)) + 
  facet_grid(. ~ cyl)

```

## Geometric objects

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = cty, y = hwy))
```

```{r}

ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = cty, y = hwy, linetype = drv))

```

multiple geom

```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy)) +
  geom_smooth(mapping = aes(x = cty, y = hwy))

```

```{r}

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()

```

add filter to show suv by smooth line

```{r}

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "suv"), se = FALSE)

```

## Statistical transformations

geom_bar()

```{r}

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = color))

? diamonds

```

stat_summary()

```{r}

ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = carat),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )

```

## Position adjustments

position = "identity", "dodge" or "fill"

```{r}

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color))

```

```{r}

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color),
           position = "dodge"
  )

```

```{r}

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color),
           alpha = 1/5, 
           position = "identity"
  )

```

position = "jitter"
  add random noise

```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy), position = "jitter")

```

## Coordinate systems

coord_flip()
  switches the x and y axes
  
coord_quickmap()
  sets the aspect ratio correctly for maps
  
coord_polar()
  uses polar coordinates

```{r}

graph <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color))

graph + coord_polar()

```
