---
title: "week10"
author: "Lu Zhang"
date: "7/28/2019"
output: html_document
---

## K-Means Clustering

Cluster

```{r}

set.seed(1234)
x <- rnorm(12, mean = rep(1:3, each = 4), sd = 0.2)
y <- rnorm(12, mean = rep(c(1, 2, 1), each = 4), sd = 0.2)
plot(x, y, col = "blue", pch = 19, cex = 2)
text(x + 0.05, y + 0.05, labels = as.character(1:12))

```


```{r}


dataFrame <- data.frame(x, y)
kmeansObj <- kmeans(dataFrame, centers = 3)
names(kmeansObj)

```

kmeans() function
  which cluster each data point got assigned to

```{r}

kmeansObj$cluster

```


## Building heatmaps from K-means solutions

```{r}

## K-means solution

set.seed(1234)
dataMatrix <- as.matrix(dataFrame)[sample(1:12), ]
kmeansObj <- kmeans(dataMatrix, centers = 3)

## make an image plot (heatmap)

par(mfrow = c(1, 2))
image(t(dataMatrix)[, nrow(dataMatrix):1], yaxt = "n", main = "Original Data")
image(t(dataMatrix)[, order(kmeansObj$cluster)], yaxt = "n", main = "Clustered Data")

```

## visualize the distance matrix

dataset

```{r}

## built-in R data set USArrests

df <- USArrests

## remove missing values

df <- na.omit(df)

df <- scale(df)
head(df)

```

get_dist

fviz_dist
  from the factoextra R package

```{r}

install.packages("factoextra")

```

fviz_dist
  visualizing a distance matrix
  
```{r}

distance <- get_dist(df)

fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))

```

## Computing k-means clustering

```{r}

k2 <- kmeans(df, centers = 2, nstart = 25)
str(k2)

k2

```

fviz_cluster
  plot the data points

illustration of the clusters

```{r}

fviz_cluster(k2, data = df)

```

standard pairwise scatter plots

```{r}

df %>%
  as_tibble() %>%
  mutate(cluster = k2$cluster,
         state = row.names(USArrests)) %>%
  ggplot(aes(UrbanPop, Murder, color = factor(cluster), label = state)) +
  geom_text()

```

3, 4, and 5 clusters

```{r}

k3 <- kmeans(df, centers = 3, nstart = 25)
k4 <- kmeans(df, centers = 4, nstart = 25)
k5 <- kmeans(df, centers = 5, nstart = 25)

# plots to compare

p1 <- fviz_cluster(k2, geom = "point", data = df) + ggtitle("k = 2")
p2 <- fviz_cluster(k3, geom = "point",  data = df) + ggtitle("k = 3")
p3 <- fviz_cluster(k4, geom = "point",  data = df) + ggtitle("k = 4")
p4 <- fviz_cluster(k5, geom = "point",  data = df) + ggtitle("k = 5")

library(gridExtra)
grid.arrange(p1, p2, p3, p4, nrow = 2)

```

