---
title: "week11"
author: "Lu Zhang"
date: "7/28/2019"
output: html_document
---

# Principal Component Analysis

two functions:

prcomp(x, center = TRUE, scale. = TRUE)
princomp(x, cor = FALSE, scores = TRUE)

## Compute the Principal Components

```{r}

mtcars.pca <- prcomp(mtcars[,c(1:7,10,11)], center = TRUE,scale. = TRUE)

summary(mtcars.pca)

```

show PCA object

```{r}

str(mtcars.pca)

```

## Plotting PCA

```{r}
library(devtools)

install_github("vqv/ggbiplot")

```

call ggbiplot on PCA

```{r}

library(ggbiplot)

ggbiplot(mtcars.pca)
```

provide an argument to ggbiplot
  labels rownames of mtcars

```{r}

ggbiplot(mtcars.pca, labels=rownames(mtcars))

```

## Interpreting the results

draw an ellipse around each group
groups argument of ggbiplot

```{r}

mtcars.country <- c(rep("Japan", 3), 
                    rep("US",4), 
                    rep("Europe", 7),
                    rep("US",3), 
                    "Europe", 
                    rep("Japan", 3), 
                    rep("US",4), 
                    rep("Europe", 3), 
                    "US", 
                    rep("Europe", 3)
                    )

ggbiplot(mtcars.pca,ellipse=TRUE,  labels=rownames(mtcars), groups=mtcars.country)

```

choices argument
  plot other components

eg. PC3 and PC4

```{r}

ggbiplot(mtcars.pca,
         ellipse=TRUE,
         choices=c(3,4),   
         labels=rownames(mtcars), 
         groups=mtcars.country)

```

## Graphical parameters with ggbiplot

circle argument

```{r}

ggbiplot(mtcars.pca,
         ellipse=TRUE,
         circle=TRUE, 
         labels=rownames(mtcars), 
         groups=mtcars.country)

```

scale the samples (obs.scale) and the variables (var.scale)

```{r}

ggbiplot(mtcars.pca,
         ellipse=TRUE,
         obs.scale = 1, 
         var.scale = 1,  
         labels=rownames(mtcars), 
         groups=mtcars.country)

```



var.axes=FALSE
  remove the arrows altogether


```{r}

ggbiplot(mtcars.pca,
         ellipse=TRUE,
         obs.scale = 1, 
         var.scale = 1,
         var.axes=FALSE,   
         labels=rownames(mtcars), 
         groups=mtcars.country)

```

## Customize ggbiplot

scale_colour_manual()
    Specify the colours for the groups
ggtitle()
    Add a title
minimal()
    Specify the minimal() theme
theme()
    Move the legend position
```{r}

ggbiplot(mtcars.pca,
         ellipse=TRUE,
         obs.scale = 1, 
         var.scale = 1,  
         labels=rownames(mtcars), 
         groups=mtcars.country) +
  scale_colour_manual(name="Origin", 
                      values= c("forest green", 
                                "red3", 
                                "dark blue"))+
  ggtitle("PCA of mtcars dataset")+
  theme_minimal()+
  theme(legend.position = "bottom")

```


## Adding a new sample

add anew sample to mtcars, creating mtcarsplus, then repeat the analysis


```{r}

spacecar <- c(1000,60,50,500,0,0.5,2.5,0,1,0,0)

mtcarsplus <- rbind(mtcars, 
                    spacecar)
mtcars.countryplus <- c(mtcars.country, 
                        "Jupiter")

mtcarsplus.pca <- prcomp(mtcarsplus[,c(1:7,10,11)], 
                         center = TRUE,
                         scale. = TRUE)

ggbiplot(mtcarsplus.pca, 
         obs.scale = 1, 
         var.scale = 1, 
         ellipse = TRUE, 
         circle = FALSE, 
         var.axes=TRUE, 
         labels=c(rownames(mtcars), "spacecar"), 
         groups=mtcars.countryplus)+
  scale_colour_manual(name="Origin", 
                      values= c("forest green", 
                                "red3", 
                                "violet", 
                                "dark blue"))+
  ggtitle("PCA of mtcars dataset, 
          with extra sample added")+
  theme_minimal()+
  theme(legend.position = "bottom")

```

## Project a new sample onto the original PCA

1. scale the values for spacecar in relation to the PCA's center (mtcars.pca$center)
2. apply the rotation of the PCA matrix to the spacecar sample
3. rbind() the projected values for spacecar to the rest of the pca$x matrix


```{r}

s.sc <- scale(t(spacecar[c(1:7,10,11)]), 
              center= mtcars.pca$center)
s.pred <- s.sc %*% mtcars.pca$rotation


mtcars.plusproj.pca <- mtcars.pca
mtcars.plusproj.pca$x <- rbind(mtcars.plusproj.pca$x, s.pred)


ggbiplot(mtcars.plusproj.pca, obs.scale = 1, 
         var.scale = 1, 
         ellipse = TRUE, 
         circle = FALSE, 
         var.axes=TRUE, 
         labels=c(rownames(mtcars), 
                  "spacecar"), 
         groups=mtcars.countryplus)+
  scale_colour_manual(name="Origin", 
                      values= c("forest green", "red3", "violet", "dark blue"))+
  ggtitle("PCA of mtcars dataset, 
          with extra sample projected")+
  theme_minimal()+
  theme(legend.position = "bottom")

```

