---
title: "week13"
author: "Lu Zhang"
date: "7/28/2019"
output: html_document
---

# Model Diagnostics

## Model Assumptions

## Checking Assumptions

use data simulated from three models

### 1 Fitted versus Residuals Plot

```{r}

sim_1 = function(sample_size = 500) {
  x = runif(n = sample_size) * 5
  y = 3 + 5 * x + rnorm(n = sample_size, mean = 0, sd = 1)
  data.frame(x, y)
}

sim_2 = function(sample_size = 500) {
  x = runif(n = sample_size) * 5
  y = 3 + 5 * x + rnorm(n = sample_size, mean = 0, sd = x)
  data.frame(x, y)
}

sim_3 = function(sample_size = 500) {
  x = runif(n = sample_size) * 5
  y = 3 + 5 * x ^ 2 + rnorm(n = sample_size, mean = 0, sd = 5)
  data.frame(x, y)
}

```


## Fitted versus Residuals Plot

useful for checking both the linearity and constant variance assumptions

```{r}

set.seed(42)
sim_data_1 = sim_1()
head(sim_data_1)

```

 fit the model and add the fitted line to a scatterplot

```{r}

plot(y ~ x, 
     data = sim_data_1, 
     col = "grey", 
     pch = 20,
     main = "Data from Model 1")
fit_1 = lm(y ~ x, 
           data = sim_data_1)
abline(fit_1, 
       col = "darkorange", 
       lwd = 3)


```

fitted versus residuals plot

```{r}

plot(fitted(fit_1), 
     resid(fit_1), 
     col = "grey", 
     pch = 20,
     xlab = "Fitted", 
     ylab = "Residuals", 
     main = "Data from Model 1")
abline(h = 0, 
       col = "darkorange", 
       lwd = 2)

```

Model 2
  non-constant variance

(the variance is larger for larger values of the predictor variable x)

```{r}

set.seed(42)
sim_data_2 = sim_2()
fit_2 = lm(y ~ x, data = sim_data_2)
plot(y ~ x, data = sim_data_2, col = "grey", pch = 20,
     main = "Data from Model 2")
abline(fit_2, col = "darkorange", lwd = 3)

```

fitted versus residuals plot 

```{r}

plot(fitted(fit_2), 
     resid(fit_2), 
     col = "grey", 
     pch = 20,
     xlab = "Fitted", 
     ylab = "Residuals", 
     main = "Data from Model 2")
abline(h = 0, 
       col = "darkorange", 
       lwd = 2)

```


Model 3 
  Y is not a linear combination of the predictors. 
  Y ~ X^2

```{r}

set.seed(42)
sim_data_3 = sim_3()
fit_3 = lm(y ~ x, 
           data = sim_data_3)
plot(y ~ x, 
     data = sim_data_3, 
     col = "grey", 
     pch = 20,
     main = "Data from Model 3")
abline(fit_3, 
       col = "darkorange", 
       lwd = 3)

```






```{r}

plot(fitted(fit_3), 
     resid(fit_3), 
     col = "grey", 
     pch = 20,
     xlab = "Fitted", 
     ylab = "Residuals", 
     main = "Data from Model 3")
abline(h = 0, 
       col = "darkorange", 
       lwd = 2)

```

### 2 Breusch-Pagan Test

```{r}

#install.packages("lmtest")
library(lmtest)

```


```{r}

bptest(fit_1)

bptest(fit_2)

bptest(fit_3)

```

### 3 Histograms

make a histogram of the residuals

```{r}

par(mfrow = c(1, 3))

hist(resid(fit_1),
     xlab   = "Residuals",
     main   = "Histogram of Residuals, fit_1",
     col    = "darkorange",
     border = "dodgerblue",
     breaks = 20)

hist(resid(fit_2),
     xlab   = "Residuals",
     main   = "Histogram of Residuals, fit_2",
     col    = "darkorange",
     border = "dodgerblue",
     breaks = 20)

hist(resid(fit_3),
     xlab   = "Residuals",
     main   = "Histogram of Residuals, fit_3",
     col    = "darkorange",
     border = "dodgerblue",
     breaks = 20)

```

### 4 Q-Q Plots

quantile-quantile plot

qqnorm() function plots the points
qqline() function adds the necessary line

```{r}

qqnorm(resid(fit_1), 
       main = "Normal Q-Q Plot, 
       fit_1", 
       col = "darkgrey")

qqline(resid(fit_1), 
       col = "dodgerblue", 
       lwd = 2)

```

Q-Q plots

```{r}

qq_plot = function(e) {

  n = length(e)
  normal_quantiles = qnorm(((1:n - 0.5) / n))
  # normal_quantiles = qnorm(((1:n) / (n + 1)))

  # plot theoretical verus observed quantiles
  plot(normal_quantiles, sort(e),
       xlab = c("Theoretical Quantiles"),
       ylab = c("Sample Quantiles"),
       col = "darkgrey")
  title("Normal Q-Q Plot")

  # calculate line through the first and third quartiles
  slope     = (quantile(e, 0.75) - quantile(e, 0.25)) / (qnorm(0.75) - qnorm(0.25))
  intercept = quantile(e, 0.25) - slope * qnorm(0.25)

  # add to existing plot
  abline(intercept, slope, lty = 2, lwd = 2, col = "dodgerblue")
}

```

qqnorm() and qqline()

```{r}

set.seed(420)
x = rnorm(100, mean = 0 , sd = 1)
par(mfrow = c(1, 2))
qqnorm(x, col = "darkgrey")
qqline(x, lty = 2, lwd = 2, col = "dodgerblue")
qq_plot(x)

# equivalent

```

perform a number of simulations
      “close to the line”
      
```{r}

# different sample sizes

par(mfrow = c(1, 3))
set.seed(420)
qq_plot(rnorm(10))
qq_plot(rnorm(25))
qq_plot(rnorm(100))

```

simulate data from a t distribution with a small degrees of freedom, for different sample sizes.

```{r}

par(mfrow = c(1, 3))
set.seed(420)
qq_plot(rt(10, df = 4))
qq_plot(rt(25, df = 4))
qq_plot(rt(100, df = 4))

```

simulate data from an exponential distribution

```{r}

par(mfrow = c(1, 3))
set.seed(420)
qq_plot(rexp(10))
qq_plot(rexp(25))
qq_plot(rexp(100))

```

create a Q-Q plot for each to assess normality of errors

```{r}

qqnorm(resid(fit_1), main = "Normal Q-Q Plot, fit_1", col = "darkgrey")
qqline(resid(fit_1), col = "dodgerblue", lwd = 2)

```

fit_1, perfect Q-Q plot,
  errors follow a normal distribution

```{r}

qqnorm(resid(fit_2), 
       main = "Normal Q-Q Plot, 
       fit_2", 
       col = "darkgrey")
qqline(resid(fit_2), 
       col = "dodgerblue", 
       lwd = 2)

```

fit_2, suspect Q-Q plot

```{r}

qqnorm(resid(fit_2), main = "Normal Q-Q Plot, fit_2", col = "darkgrey")
qqline(resid(fit_2), col = "dodgerblue", lwd = 2)

```

fit_3, suspect Q-Q plot

```{r}

qqnorm(resid(fit_3), main = "Normal Q-Q Plot, fit_3", col = "darkgrey")
qqline(resid(fit_3), col = "dodgerblue", lwd = 2)

```


### 5 Shapiro-Wilk Test

formal testing
shapiro.test()

```{r}

set.seed(42)
shapiro.test(rnorm(25))

## W = 0.71164, p-value = 1.05e-05

shapiro.test(rexp(25))

## W = 0.99858, p-value = 0.9622

shapiro.test(resid(fit_3))

## W = 0.97643, p-value = 3.231e-07
```

# Outliers

resid() for residual
rstandard() for Standardized residual